Chemical tagging can work: \\ Identification of stellar phase-space structures \\ purely by chemical-abundance similarity

### 2016-01-20 papers/clusters.tex
Subtle emphasis on the model size
Surely the [logarithmic] brackets around these scalings was unintended, right?? @davidwhogg

Our analysis draws on the spectra of \totalnumber\ giant stars ($\logg
< 3.9$) from \apogee\ \acronym{DR12} \citep{dr12}, with no warning flags
set in the \apogee\ \aspcap\ \citep{aspcap} pipeline
reductions.
We re-analyze these \emph{spectra} using \thecannon, because this
deliver stellar abundance labels of higher precision, especially for
stars of $S/N \le 150$.
The details about how we select, reduce, and analyze the \apogee\ data
are given in full detail in the companion papers \citep{casey16,
  ness16}, and we only summarize briefly here.
We re-derived 17 labels ($\teff$, $\logg$, and 15 abundances
referenced to H) from the \apogee\ \acronym{DR12} spectra. 
For the training step of \thecannon, 
12,681 red-giant stars with spectral $S/N \ge 200$ were used.
For this small fraction of the sample with the highest $S/N$, the
labels provided by \aspcap\ provide consistent, relatively low-scatter, sensible
abundance-space measurements.
The training step fixes a spectral model
that predicts the normalized spectrum as a quadratic function of all
labels. In the second stage---the test step---each unlabelled star is
used to establish a single-star likelihood function for its labels,
holding the model coefficients fixed.
\thecannon\ finds the labels that minimize the single-star likelihood
function for each test-set star.
This optimization is not convex, but it is embarrassingly parallel,
and therefore fast. The test
step for 150,000 spectra takes less than 30~minutes on a small 
research cluster in Cambridge.

@@

Another issue with k-means is that it effectively uses metric
distances in the $D$-space; it presumes Euclidean isotropy.
We choose here to work in the hydrogen-normalized abundance space, the
space of [C/H], [N/H], [O/H], [Na/H], [Mg/H], [Al/H], [Si/H], [S/H],
[K/H], [Ca/H], [Ti/H], [V/H], [Mn/H], [Fe/H], [Ni/H].
But in addition to this, we re-scale these by approximate typical measurement
precisions obtained by \thecannon\ before running k-means.
These scalings were
  [C/H]/[0.041],
  [N/H]/[0.044],
  [O/H]/[0.037],
  [Na/H]/[0.111],
  [Mg/H]/[0.032],
  [Al/H]/[0.055],
  [Si/H]/[0.041],
  [S/H]/[0.054],
  [K/H]/[0.069],
  [Ca/H]/[0.043],
  [Ti/H]/[0.072],
  [V/H]/[0.146],
  [Mn/H]/[0.041],
  [Fe/H]/[0.019], and
  [Ni/H]/[0.034].
This scaling makes the space close to isotropic in measurement uncertainty or
observational precision.
Importantly we use \emph{only} abundance-space information, and no
positional or velocity information (nor $\teff$ nor $\logg$ nor any
targeting or observational meta-data) as input to the clustering
algorithm.

>>>>

Our analysis draws on the spectra of \totalnumber\ giant stars ($\logg
< 3.9$) from \apogee\ \acronym{DR12} \citep{dr12}, with no warning flags
set in the \apogee\ \aspcap\ \citep{aspcap} pipeline
reductions.
We re-analyze these \emph{spectra} using \thecannon, because this
deliver stellar abundance labels of higher precision, especially for
stars of $S/N \le 150$.
The details about how we select, reduce, and analyze the \apogee\ data
are given in full detail in the companion papers \citep{casey16,
  ness16}, and we only summarize briefly here.
We re-derived 17 labels ($\teff$, $\logg$, and 15 abundances
referenced to H) from the \apogee\ \acronym{DR12} spectra. 
For the training step of \thecannon, 
12,681 red-giant stars with spectral $S/N \ge 200$ were used.
For this small fraction of the sample with the highest $S/N$, the
labels provided by \aspcap\ provide consistent, relatively low-scatter, sensible
abundance-space measurements.
The training step fixes a spectral model
that predicts the normalized spectrum as a quadratic function of all
labels ($\sim 1.5\times10^6$ model coefficients). In the second stage---the test step---each unlabelled star is
used to establish a single-star likelihood function for its labels,
holding the model coefficients fixed.
\thecannon\ finds the labels that minimize the single-star likelihood
function for each test-set star.
This optimization is not convex, but it is embarrassingly parallel,
and therefore fast. The test
step for 150,000 spectra takes less than 30~minutes on a small 
research cluster in Cambridge.

@@

Another issue with k-means is that it effectively uses metric
distances in the $D$-space; it presumes Euclidean isotropy.
We choose here to work in the hydrogen-normalized abundance space, the
space of [C/H], [N/H], [O/H], [Na/H], [Mg/H], [Al/H], [Si/H], [S/H],
[K/H], [Ca/H], [Ti/H], [V/H], [Mn/H], [Fe/H], [Ni/H].
But in addition to this, we re-scale these by approximate typical measurement
precisions obtained by \thecannon\ before running k-means.
These scalings were
  [C/H]/0.041,
  [N/H]/0.044,
  [O/H]/0.037,
  [Na/H]/0.111,
  [Mg/H]/0.032,
  [Al/H]/0.055,
  [Si/H]/0.041,
  [S/H]/0.054,
  [K/H]/0.069,
  [Ca/H]/0.043,
  [Ti/H]/0.072,
  [V/H]/0.146,
  [Mn/H]/0.041,
  [Fe/H]/0.019, and
  [Ni/H]/0.034.
This scaling makes the space close to isotropic in measurement uncertainty or
observational precision.
Importantly we use \emph{only} abundance-space information, and no
positional or velocity information (nor $\teff$ nor $\logg$ nor any
targeting or observational meta-data) as input to the clustering
algorithm.
